Technical Report

### 2018-06-14 fully-abstract-comp/technical-report.tex
Suggestion for simplification of the reasonability condition.

\dominique{The below is an experimental reformulation of the reasonability requirement for the FTLR and full abstraction formulation.  All remarks welcome..
  The main change is that it now says that after a call, the post-return code is also expected to behave reasonably under reasonable assumptions. 
}
\dominique{TODO: better name than ``reasonable'' for all of this.}
Take a block of code memory $\mscode$, a set of trusted addresses $\ta$ and sets of return pointer and closure seals $\sigrets$ and $\sigcloss$.
We define that a word $w$ is reasonable up to $n$ steps in memory $\ms$ if the following implications hold.
\begin{itemize}
\item If $w = \seal{\sigma_\baddr,\sigma_\eaddr,\_}$, then $[\sigma_\baddr,\sigma_\eaddr] \mathrel{\#} (\sigrets \cup \sigcloss)$
\item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ then $[\baddr,\eaddr] \# \ta$
\item If $w = \sealed{\sigma,\vsc}$ and $\sigma \not\in (\sigrets\cup \sigcloss)$ then $\vsc$ is reasonable.
\item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\perm \in \readAllowed{}$ and $n > 0$, then $\ms(\aaddr)$ is readable up to $n - 1$ steps for all $\aaddr \in [\baddr,\eaddr]$
\end{itemize}

We say that an executable capability $c = ((\perm,\normal),\baddr,\eaddr,\aaddr)$ and data capability $c_{\mathrm{data}}$ behaves reasonably with stack assumption $\ms_\stk'$ up to $n$ steps if the following implication holds:

For any $\Phi$ such that
  \begin{itemize}
  \item $\Phi.\reg(\pcreg) = c$
  \item $\Phi.\reg(\rdata) = c_{\mathrm{data}}$
  \item $\Phi.\reg(r)$ is reasonable up to $n$ steps in memory $\Phi.\mem$ for all $r \not\in \{\pcreg,\rdata\}$
  \item $\Phi.\ms_\stk$ is reasonable up to $n-1$ steps in memory $\Phi.\mem$ for all $a$
  \item $\Phi.\ms_\stk|_{\dom(\ms_\stk')} = \ms_\stk'$
  \item $\Phi.\mem$, $\Phi.\ms_\stk$ and $\Phi.\stk$ are all disjoint
  \end{itemize}
  We have the following implications:
  \begin{itemize}
  \item \emph{Guarantee stack base address before call} If
    \begin{itemize}
    \item $\Phi \nstep[i]{\ta,\stkb} \Phi'$
    \item $\Phi'$ points to $\src{\scall{\offpc,\offsigma}{r_1}{r_2}}$ in $\ta$ for some $\src{r_1}$ and $\src{r_2}$
    \end{itemize}
    Then all of the following hold:
    \begin{itemize}
    \item $\src{\Phi'}(\src{r_\stk}) = \src{\stkptr{\_,\stkb,\eaddr_\stk,\aaddr_\stk}}$
    \item $r_1 \neq \rtmp{1}$
    \item $\Phi'(\pcreg) + \calllen$ and $\Phi'(\rstk)$ behave reasonably with stack assumption $\Phi.\ms_\stk|_{[\eaddr_\stk,\aaddr_\stk]}$ up to $n-i$ steps
    \end{itemize}
  \item \emph{Use return seals only for calls} If
    \begin{itemize}
    \item $\Phi \nstep[i]{\ta,\stkb} \Phi'$
    \item $\src{\Phi'}$ points to $\src{\tcseal{r_1}{r_2}}$ in $\ta$ and $\Phi'(r_1) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$
    \end{itemize}
    Then one of the following holds:
    \begin{itemize}
    \item $\src{\Phi'}$ is inside $\src{\scall{\offpc,\offsigma}{r_1'}{r_2'}}$ and $\sigma \in \sigrets$ 
    \item $\sigma \in \sigcloss$.
    \end{itemize}
  \item \emph{Don't store private stuff...} If
    \begin{itemize}
    \item $\Phi \nstep[i]{\ta,\stkb} \Phi'$
    \item $\src{\Phi'}$ points to $\src{\tstore{r_1}{r_2}}$ in $\ta$, then 
    \end{itemize}
    Then $\Phi.\reg(r_2)$ is reasonable in memory $\Phi'.\mem$ up to $n$ steps
  \item \emph{Don't leak private stuff...} If
    \begin{itemize}
    \item $\Phi \nstep[i]{\ta,\stkb} \Phi' \step \Phi''$
    \item $\src{\Phi'}$ is in $\ta$
    \item $\src{\Phi''}$ is not in $\ta$
    \end{itemize}
    Then $\Phi.\reg(r)$ is reasonable in memory $\Phi'.\mem$ up to $n$ steps for all $r$ (including $\pcreg$)
  \end{itemize}

We say that a component $(\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss,A_\linear)$ is reasonable if the following hold:
For all $\sealed{\sigma,\vsc},\sealed{\sigma,\vsc'} \in \overline{c_{\mathrm{export}}}$, with $\exec{\vsc}$ and $\nonExec{\vsc'}$, we have that $\vsc$ and $\vsc'$ behave reasonably with stack assumption $\emptyset$ up to any number of steps $n$.

>>>>

\dominique{The below is an experimental reformulation of the reasonability requirement for the FTLR and full abstraction formulation.  All remarks welcome..
  The main change is that it now says that after a call, the post-return code is also expected to behave reasonably under reasonable assumptions. 
}
\dominique{TODO: better name than ``reasonable'' for all of this.}
Take a block of code memory $\mscode$, a set of trusted addresses $\ta$ and sets of return pointer and closure seals $\sigrets$ and $\sigcloss$.
We define that a word $w$ is reasonable up to $n$ steps in memory $\ms$ if the following implications hold.
\begin{itemize}
\item If $w = \seal{\sigma_\baddr,\sigma_\eaddr,\_}$, then $[\sigma_\baddr,\sigma_\eaddr] \mathrel{\#} (\sigrets \cup \sigcloss)$
\item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ then $[\baddr,\eaddr] \# \ta$
\item If $w = \sealed{\sigma,\vsc}$ and $\sigma \not\in (\sigrets\cup \sigcloss)$ then $\vsc$ is reasonable.
\item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\perm \in \readAllowed{}$ and $n > 0$, then $\ms(\aaddr)$ is reasonable up to $n - 1$ steps for all $\aaddr \in ([\baddr,\eaddr] \setminus \ta)$
\end{itemize}

We say that an executable capability $c = ((\perm,\normal),\baddr,\eaddr,\aaddr)$ behaves reasonably up to $n$ steps if the following implication holds:

For any $\Phi$ such that
  \begin{itemize}
  \item $\Phi.\reg(\pcreg) = c$
  \item $\Phi.\reg(r)$ is reasonable up to $n$ steps in memory $\Phi.\mem$ for all $r \not\in \{\pcreg\}$
  \item $\Phi.\ms_\stk$ is reasonable up to $n-1$ steps in memory $\Phi.\mem$ for all $a$
  \item $\Phi.\mem$, $\Phi.\ms_\stk$ and $\Phi.\stk$ are all disjoint
  \end{itemize}
  We have the following implications:
  \begin{itemize}
  \item \emph{Guarantee stack base address before call} If
    \begin{itemize}
    \item $\Phi \nstep[i]{\ta,\stkb} \Phi'$
    \item $\Phi'$ points to $\src{\scall{\offpc,\offsigma}{r_1}{r_2}}$ in $\ta$ for some $\src{r_1}$ and $\src{r_2}$
    \end{itemize}
    Then all of the following hold:
    \begin{itemize}
    \item $\src{\Phi'}(\src{r_\stk}) = \src{\stkptr{\_,\stkb,\eaddr_\stk,\aaddr_\stk}}$
    \item $r_1 \neq \rtmp{1}$
    \item $\Phi'(\pcreg) + \calllen$ and $\Phi'(\rstk)$ behaves reasonably up to $n-i$ steps
    \end{itemize}
  \item \emph{Use return seals only for calls} If
    \begin{itemize}
    \item $\Phi \nstep[i]{\ta,\stkb} \Phi'$
    \item $\src{\Phi'}$ points to $\src{\tcseal{r_1}{r_2}}$ in $\ta$ and $\Phi'(r_1) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$
    \end{itemize}
    Then one of the following holds:
    \begin{itemize}
    \item $\src{\Phi'}$ is inside $\src{\scall{\offpc,\offsigma}{r_1'}{r_2'}}$ and $\sigma \in \sigrets$ 
    \item $\sigma \in \sigcloss$.
    \end{itemize}
  \item \emph{Don't store private stuff...} If
    \begin{itemize}
    \item $\Phi \nstep[i]{\ta,\stkb} \Phi'$
    \item $\src{\Phi'}$ points to $\src{\tstore{r_1}{r_2}}$ in $\ta$, then 
    \end{itemize}
    Then $\Phi.\reg(r_2)$ is reasonable in memory $\Phi'.\mem$ up to $n$ steps
  \item \emph{Don't leak private stuff...} If
    \begin{itemize}
    \item $\Phi \nstep[i]{\ta,\stkb} \Phi' \step \Phi''$
    \item $\src{\Phi'}$ is in $\ta$
    \item $\src{\Phi''}$ is not in $\ta$
    \end{itemize}
    Then $\Phi.\reg(r)$ is reasonable in memory $\Phi'.\mem$ up to $n$ steps for all $r$ (including $\pcreg$)
  \end{itemize}

We say that a component $(\mscode,\msdata,\overline{c_{\mathrm{import}}},\overline{c_{\mathrm{export}}},\sigrets,\sigcloss,A_\linear)$ is reasonable if the following hold:
For all $\sealed{\sigma,\vsc},\sealed{\sigma,\vsc'} \in \overline{c_{\mathrm{export}}}$, with $\exec{\vsc}$ and $\nonExec{\vsc'}$, we have that $\vsc$ and $\vsc'$ behave reasonably with stack assumption $\emptyset$ up to any number of steps $n$.
\lau{This need to change with the proposed simplification of the reasonability update.}
